---
title: "Statistical Inference, Part 1"
author: "Aditya Salapaka"
date: "Friday 20 March 2015"
output: html_document
---

# A Basic Analysis of the Exponential Function in R

## Overview
An exponential distribution with lambda = 0.2 and 40 observations was simulated
1000 times, and its mean and standard deviation were calculated. These were 
compared with the theoretical mean and standard deviation. A histogram of the 
simulated dataset was also created, and a normal distribution curve was 
superimposed on it to prove that the simulated dataset was approximately normal.

## The Exponential Distribution
The exponential function is defined as:
$$ f(x,λ) = e^{-λx} $$
where λ is the rate parameter. 

The mean of this distribution is ${1}/{λ}$.

## Simulations
The exponential distribution can be generated in R by the `rexp()` function. I
generated an exponential distribution with 40 observations and λ = 0.2 and took
its mean. I ran this simulation 1000 times and created a data frame with the
averages of all these distributions. The seed for the pseudo-random generator
was set at 850.

```{r}
lambda <- 0.2
n <- 40
sims <- 1:1000

set.seed(850)

mns <- data.frame(x = sapply(sims, function(x){
        mean(rexp(n, lambda))
}))
```

## Sample Mean vs Theoretical Mean
The sample mean is `r mean(mns$x)`, which is close to the theoretical mean 
`r 1/lambda`.

## Sample variance vs Theoretical Variance
The theoretical variance of this distribution is calculated by applying the
central limit theorem. The variacne is $\frac{1/λ}{\sqrt{n}}$
The sample variance is `r var(mns$x)`, which is close to the theoretical
variance `r ((1/lambda)/sqrt(n))^2`.

## Distribution
A histogram of the sample data superimposed with a normal distribution curve is
shown. It is clear that the distribution is approxmimately normal.

```{r}
require("ggplot2")
g <- ggplot(mns, aes(x = mns$x))
p <- g + geom_histogram(aes(y = ..density..), binwidth = 0.4, fill = "white", color = "black") + 
        stat_function(fun = dnorm, arg = list(mean = 5, sd = sd(mns$x)), colour = "red") + 
        xlab("Means") + ylab("Density")
print(p)
```