---
title: "Statistical Inference, Part 2"
author: "Aditya Salapaka"
date: "Sunday 22 March 2015"
output: pdf_document
---

# Hypothesis Testing on the ToothGrowth dataset

## Overivew
The ToothGrowth dataset measures the length of teeth in each of the 10 guinea
pigs given three doses levels of Vitamin C (0.5, 1 and 2mg) with each of two 
delivery methods (orange juice and ascorbic acid). Via this document, I will
try to understand the effect of Vitamin C of different dosages via different
administration methods on the length of the teeth of these guinea pigs.

## Data
The ToothGrowth dataset is first loaded and observed.

```{r}
data(ToothGrowth)
head(ToothGrowth)
```

For a basic summary of the dataset:
```{r}
str(ToothGrowth)
```

A plot of the dataset is as follows:
```{r warning=FALSE}
require("ggplot2")
g <- ggplot(ToothGrowth, aes(dose, len))
p <- g + geom_point(aes(x = dose, y = len, colour = supp)) + 
        geom_smooth(aes(colour = supp), method = "loess")
print(p)
```

The red line `OJ` represents the Orange Juice delivery method and blue line `VC`
represents the Ascorbic Acid delivery method. It is clear that while there 
exists a significant difference between the lengths of teeth in lower dosages 
(with `OJ` having a greater impact), this difference is not as pronounced at
higher dosages. I will attempt to verify this by hypothesis testing.

## Assumptions
Here, I assume that the populations have equal variances and the sample means
are normally distributed. 

## Null Hypothesis
There is no difference between the length of teeth when Vitamin C is delivered
via Orange Juice or Ascorbic Acid. 

I separated the ToothGrowth into two data frames `oj` and `vc`, which groups
them based on the method of delivery.

```{r}
vc <- ToothGrowth[ToothGrowth$supp == 'VC',]
oj <- ToothGrowth[ToothGrowth$supp == 'OJ',]
```

## Hypothesis Testing
Here, I will perform the paired t test with, which has a default confidence 
level of 95%.

For a dosage of 0.5mg:

```{r}
t.test(vc$len[vc$dose == 0.5], oj$len[oj$dose == 0.5], paired = TRUE, 
       var.equal = TRUE)
```

The 95% confidence values gives the interval -9.23 to -1.26, and the mean of 
the differences is -5.25. The p-value is 0.015, which is lesser than 0.05. So
we can reject the null hypothesis for a dosage of 0.5mg. Therefore, different 
methods of administration does make a difference in the length of teeth of 
guinea pigs, for a dosage of 0.5mg.

For a dosage of 1mg:

```{r}
t.test(vc$len[vc$dose == 1], oj$len[oj$dose == 1], paired = TRUE, 
       var.equal = TRUE)
```

The 95% confidence values gives the interval -9.91 to -1.95, and the mean of 
the differences is -5.93. The p-value is 0.008, which is lesser than 0.01. So
we can reject the null hypothesis for a dosage of 1mg. Therefore, different 
methods of administration does make a difference in the length of teeth of 
guinea pigs, for a dosage of 1mg.

For a dosage of 2mg:

```{r}
t.test(vc$len[vc$dose == 2], oj$len[oj$dose == 2], paired = TRUE, 
       var.equal = TRUE)
```

The 95% confidence values gives the interval -4.16 to -4.32, and the mean of 
the differences is -0.08. The p-value is 0.967, which is greater than 0.05. So
we cannot reject the null hypothesis for a dosage of 2mg. Therefore, different 
methods of administration does not make a difference in the length of teeth of 
guinea pigs, for a dosage of 2mg.

## Conclusion
From the hypothesis test, we have concluded that for dosages of 0.5mg and 1mg,
the method of delivery affects the length of teeth in guinea pigs. However, this
does not hold true for 2mg.